---
title: "fish new"
output: html_document
---

```{r}
library(olsrr)
library(dummies) 
library(MASS)
library(ggplot2)

fish <- read.csv("Fish.csv",header = T)
head(fish)
names(fish)[1] <-  "Species" # adjust variable name
# a: description of the variables
summary(fish)
# Species is a categorical variable with 7 possible values

# b: bar-plot
ggplot(data= fish,aes(x=Species)) + geom_bar()

# c: Yes. Species should be broken into 7 binary variables,
#     only 6 of which should be included in the model. The omitted one is the "basis" for beta_0.

```
```{r}

# d: build a model

dum <- dummy(fish$Species, sep = ".")
new_fish <- cbind(fish[,-1], dum)
model <- lm(Weight ~. , data = new_fish[,-ncol(new_fish)])
summary(model) 

# e: Species.Pike and Species.Smelt, this indicates that fish from those species have different weights than from other species, given the same
# values of other parameters. 
# Note that Whitefish is our baseline hence omittied. Note that if you chose other basis, you could have got other significant Species. 
# Length 1 is also significant. Length2 is not for alpha 0.05 but seems to be important since it is for alpha=0.083.

# f: R^2: 0.9361. 93% of the variability in Y is explained by X.

# g,h,i: multicollinearity analysis.
coll_ans <- ols_coll_diag(model)

coll_ans$vif_t
# most toleranc and VIF values indicate multicollinearity.

round(cor(new_fish),2)
# length1, length2, length3 are extremely correlated. They are also very correlated with Height and Width.

round(coll_ans$eig_cindex,3)
# length2 is almost completly explained by the the factor with the largest condition index. Length 1 is also at large explained by it.
# width, height are largly explained by the factor with 54.49 condition index. length3 by the factor with 347.65 condition index.
# all those indicate multicollinearity.

```
```{r}

# since length1, length2, length3 are so strongly correlated it is almost a state of complete multicollinearity. We will try to remove 2 of them.
model2 <- lm(Weight ~  Height + Length1 +  Width + 
               Species.Bream + Species.Parkki + Species.Perch + Species.Pike + Species.Roach + Species.Smelt, data = new_fish)

# R^2 is 0.9309, almost no change. This is because almost all the information in length2 and length3 is already included in the model by adding length1.

coll_ans2 <- ols_coll_diag(model2)
coll_ans2$vif_t
round(coll_ans2$eig_cindex, 3)
# Still all the metrics indicate that there is a problem.
# We can see that length1, height and some of the species are still part of an approximate linear combination.
# You should have done another mid step, but we'll remove both Height and Width.

model3 <- lm(Weight ~ Length1 + Species.Bream + Species.Parkki + Species.Perch + Species.Pike + Species.Roach + Species.Smelt, data = new_fish)
# R^2 still hasn't changed (96.83)

coll_ans3 <- ols_coll_diag(model3)
coll_ans3$vif_t
round(coll_ans3$eig_cindex, 3)

# looks much better now.
# now, after we've learned some model fit metrics (such as BIC, AIC) we can comapre models acording
# to those metric as well
# try repeat the previous analysis and looking at AIC and R^2-adjusted 

summary(model) 
summary(model2) 
summary(model3) 

```
```{r}

# j: error analysis
e <- resid(model3)

#histogram of residuals
hist(e, breaks=20) # looks skewed

# QQ Plot
qqnorm(e)
qqline(e) #confirms skew

# Standartized residuals vs. prediction
r <- stdres(model3)
y_hat <- model3$fitted.values
plot(r, y_hat, xlab = "Standartized Residuals", ylab = "Prediction", cex=0.5)

# Standartized residuals vs. Length1
plot(r, new_fish$Length1, xlab = "Standartized Residuals", ylab = "Length1", cex=0.5)

# Seems to be a connection between r and y_hat, caused by Length1. Higher values of Length1 have higher residual variance. This looks like heteroscedasticity.

```
```{r}

# k: Interactions

# create all interactions (you might have used only part of them)
len_Bream <- new_fish$Length1 * new_fish$Species.Bream
len_Parkki <- new_fish$Length1 * new_fish$Species.Parkki
len_Perch <- new_fish$Length1 * new_fish$Species.Perch
len_Pike <- new_fish$Length1 * new_fish$Species.Pike
len_Roach <- new_fish$Length1 * new_fish$Species.Roach
len_Smelt <- new_fish$Length1 * new_fish$Species.Smelt

interact_fish <- data.frame(cbind(new_fish$Weight, new_fish$Length1, new_fish$Species.Bream, new_fish$Species.Parkki, new_fish$Species.Perch, new_fish$Species.Pike, new_fish$Species.Roach, new_fish$Species.Smelt,
                       len_Bream, len_Parkki, len_Perch, len_Pike, len_Roach, len_Smelt))
colnames(interact_fish)<- c("Weight", "Length1", "Species.Bream", "Species.Parkki", "Species.Perch", "Species.Pike", "Species.Roach", "Species.Smelt", "len_Bream", "len_Parkki", "len_Perch", "len_Pike", "len_Roach", "len_Smelt")
model4 <-  lm(Weight ~., data = interact_fish)
summary(model4)

# re-examine residuals

# Standartized residuals vs. prediction
r <- stdres(model4)
y_hat <- model4$fitted.values
plot(r, y_hat, xlab = "Standartized Residuals", ylab = "Prediction", cex=0.5)

# Standartized residuals vs. Length1
plot(r, interact_fish$Length1, xlab = "Standartized Residuals", ylab = "Length1", cex=0.5)

# looks significantly better.
